# Hunting for attractors

This work is a direct reuse of results from Sprott (1993)^1, Bourke [link](http://paulbourke.net/fractals/sprott/) and the R implementation of quadratic map functions from Lindberg [link](https://blog.k2h.se/post/hunting-for-attractors/).
This code sample parameter space, filter aesthetic functions based on different criterias (Lyapunov exponent, summary of point density distribution, correlation) and display them in a grid.

1. Sprott, J. C. (1993). Automatic generation of strange attractors. Computers & Graphics, 17(3), 325-332.

```{r setup}
library(tidyverse)
library(future)
library(furrr)
library(cowplot)

plan(multisession)

# options
options(dplyr.summarise.inform=FALSE)
update_geom_defaults("point", list(size = 0.1, stroke = 0, shape = 16))

# functions
source("R/attractors_functions.r")

```


```{r sample, eval=FALSE}

set.seed(1)

n_total <- 5e5

# generate a set of unique 12 letters sequences from a set of 25 letters (25^12 set)
# get parameters corresponding to sampled sequence
table_set <- tibble(pattern = 1:n_total) %>% 
  mutate(name = map(pattern, ~ sample_sequence(set = LETTERS[1:25], length = 12))) %>% 
  distinct(name, .keep_all = TRUE) %>% 
  mutate(p = map(name, ~ get_parameters(string = .))) 

# filter for functions that do not converge into a single point (25/1e4 @ 1000)
table_set <- table_set %>% 
  mutate(L = future_map_dbl(p, ~ L(quadratic_map, ., 0, 0, iterations = 1000))) %>%
  filter(L > 0) %>% unnest(name)

# compute xy data and point density as an aesthetic metric. (28 / 1000 @ 5000)
# filter diverging functions 
table_set <- table_set %>%
  mutate(xy = future_map(p, ~ iterate(quadratic_map, ., 0, 0, iterations = 5000))) %>% 
  filter(map_lgl(xy, ~ with(., all(abs(x) + abs(y) < 1e7)))) %>% 
  mutate(density = map(xy, density_metric, gridsize = 20)) %>% 
  mutate(r = map_dbl(xy, ~ with(., cor(x,y)))) %>% 
  select(-xy) %>% unnest(density)

# export
saveRDS(table_set, file="data/attractor_parameter_set_discrete.rds", compress="gzip")

```

```{r plot, eval=FALSE}

# plot 5 cm square plot in A2 format (40 x 60) n = 8*12 = 96, quality=90
# plot 5 cm square plot in A1 format (60 x 80) n = 12*16 = 192

# load results from previously intensive sampling
table_set <- read_rds("data/attractor_parameter_set_discrete.rds")

# filter functions (~ 0.5 % of sampled:aesthetic)
table_collection <- table_set %>%
  filter(between(L, 0.1, 0.4)) %>% 
  filter(d > 0.5) %>% 
  filter(cv < 1) %>% 
  filter(m < 0.5) %>% 
  filter(abs(r) < 0.8)

# compute quadratic map at high resolution for a random subset of functions
# post-process xy data (filter early iterations, rescaling)
# do two subsetting because some rare functions diverge for large iterations
set.seed(1)

data_collection <- table_collection %>%
  slice_sample(n = 100) %>% 
  mutate(xy = future_map(p, ~ iterate(quadratic_map, ., 0, 0, iterations = 4e5))) %>% 
  filter(map_lgl(xy, ~ with(., all(abs(x) + abs(y) < 1e7)))) %>% 
  mutate(xy = map(xy, ~ normalize_xy(.))) %>% 
  slice_sample(n = 96)

# compute plots for selected functions
data_collection <- data_collection %>%
  mutate(plot = map(xy, ~ render_plot(., size = 0.1))) %>% 
  arrange(name)

plot_collection <- plot_grid(
  plotlist = data_collection$plot, ncol = 8, scale = 0.7,
  labels = str_to_title(data_collection$name),
  label_x = 0.2, label_y = 0.15, label_size = 7,
  label_fontface = "italic", label_fontfamily = "NewCenturySchoolbook")

ggsave(
  plot_collection,
  file = "./R/figures/attractor/set_collection_01.png",
  dpi=300, width=400, height=600, scale=1, units="mm")

```

```{r explore, eval=FALSE}

# quick plot
plot_glance <- table_set %>% 
  filter(L > 0.1, L < 0.4) %>% 
  filter(d > 0.4) %>% 
  filter(cv < 1) %>% 
  filter(m < 0.5) %>% 
  slice_sample(n = 10) %>% 
  mutate(xy = future_map(p, ~ iterate(quadratic_map, ., 0, 0, iterations = 1e4))) %>% 
  mutate(xy = map(xy, ~ normalize_xy(.))) %>% 
  mutate(r = map_dbl(xy, ~ with(., cor(x,y)))) %>% 
  unnest(xy) %>% 
  ggplot(aes(x, y)) +
  geom_point(size = 0.3, alpha = 0.1) +
  facet_wrap(vars(name)) + coord_equal() + theme_void() 

# plot positions of aesthetic functions
plot_solution <- table_set %>%
  filter(L > 0.1) %>%
  ggplot(aes(x = d, y = cv)) +
  stat_density_2d(geom = "raster", aes(fill = after_stat(density)), contour = FALSE) +
  scale_fill_viridis_c()

# plot from letter code (e.g. WKDTLOOGQEHP FIRCDERRPVLD MCRBIPOPHTBN GLXOESFTTPSV QFFVSLMJJGCR)
plot_single <- iterate(
  quadratic_map, get_parameters(string = "SUMCUGEHROAQ"),
  0.05, 0.05, iterations = 1E5) %>%
  render_plot()

```

