# use open-data digital elevation model (ESRI ASCII Raster)

```{r setup}
library(tidyverse)
library(stars)

knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)

# functions
source("R/paper_functions.r")
source("R/gis_functions.r")
```

## Outline

Digital Elevation Model (DEM) represents elevation as function of latitude and longitude. These illustrations are made of multiple lines of elevation as a function of longitude, for discrete latitude values. Overplotting is avoided by checking for a minimal distance between two lines. These three examples explore the use of :

* a coarse and easy source for elevation data ([elevatr](https://github.com/jhollist/elevatr) package), for a broad view (~ 450 km)
* a precise DEM at a the scale of a specific valley (Ossau, FR, ~ 20 km) using IGN 5m RGE ALTI DEM ([link](https://geoservices.ign.fr/documentation/diffusion/telechargement-donnees-libres.html#rge-alti-5-m)) 
* a focus on a narrow zone (Loudenvielle, FR, ~ 3 km), with a fine initial sampling resolution for the elevation (30 cm, as reported [here](https://www.geoportail.gouv.fr/carte?c=2.4206324117987448,48.81510668542717&z=10&l0=GEOGRAPHICALGRIDSYSTEMS.MAPS.SCAN-EXPRESS.STANDARD::GEOPORTAIL:OGC:WMTS(1)&l1=ELEVATIONGRIDCOVERAGE.HIGHRES.QUALITY::GEOPORTAIL:OGC:WMTS(1)&permalink=yes)). 


```{r plot_pyrenees, eval=FALSE}

# define location and bounding box
coord = c(lat=43.231196, lon = 0) # Tarbes
z_scale = 2 # ratio of elevation:cell unit

# convert DEM to a dataframe and subset based on bounding box
point <- data.frame(lon = coord["lon"], lat = coord["lat"]) %>% 
  st_as_sf(coords = c("lon", "lat")) %>% st_set_crs(4326) %>% st_transform(crs = 2154) 

box <- buffer_rectangle(
  point, size = 450E3, y_shift = - 60E3, x_shift = 30E3,
  orientation="h", ratio = 2)

# use AWS terrain tile (30m)
dem <- elevatr::get_elev_raster(point, z = 7) 

dem_subset <- dem %>% st_as_stars() %>% 
  st_crop(box) %>% as_tibble() %>%
  select(x,y, z = 3) %>% 
  mutate(x = x, y = y, z = z * z_scale)

plot_ridge <- dem_subset %>%
  render_ridge(
    n_ridges = 200, n_threshold = 5,
    z_shift = 1000, z_threshold = 100) 

ggsave(
  plot_ridge, file = "R/figures/gis/ridge_pyrenee_00.png",
  dpi=300, width=format$a3[2], height = format$a3[2]/ratio$uni ,
  scale=1, units="mm")

```

```{r plot_ossau, eval=FALSE}


# define location and bounding box
coord = c(lat=42.9136778, lon=-0.4015139) # Cezy

# use RGE ALTI 5m DEM
dem <- coord %>% read_dem(buffer = 12000, dep="64", source = "rge_alti")

# convert DEM to a dataframe and subset based on bounding box
point <- data.frame(lon = coord["lon"], lat = coord["lat"]) %>% 
  st_as_sf(coords = c("lon", "lat")) %>% st_set_crs(4326) %>% st_transform(crs = 2154) 

box <- buffer_rectangle(point, size = 22E3, y_shift = 1E3, orientation="v")

dem_subset <- dem %>% st_crop(box) %>% 
  as_tibble() %>% mutate(x = -x, y = -y)

plot_ridge <- dem_subset %>%
  render_ridge(
    n_ridges = 300, n_drop = 10, n_threshold = 15,
    z_shift = 10, z_threshold = 8)

ggsave(
  plot_ridge, file = "R/figures/gis/ridge_ossau_01.png",
  dpi=300, width=format$a3[2], height=format$a3[2]/ratio$wide, scale=1, units="mm")

```

```{r plot_loudenvielle}

# Test ridge line rendering with true HR DEM data (5m, 30cm source)

coord = c(lat=42.744617, lon = 0.410614) # Loudenvielle
z_scale = 1 # ratio of elevation:cell unit

# convert DEM to a dataframe and subset based on bounding box
point <- data.frame(lon = coord["lon"], lat = coord["lat"]) %>% 
  st_as_sf(coords = c("lon", "lat")) %>% st_set_crs(4326) %>% st_transform(crs = 2154) 

# RGE ALTI (5m)
dem <- coord %>% read_dem(buffer = 5000, dep="65", source = "rge_alti")

box <- buffer_rectangle(point, size = 3E3, orientation="v")

dem_subset <- dem %>% st_as_stars() %>% 
  st_crop(box) %>% as_tibble() %>%
  select(x0=x, y0=y, z0 = 3) %>% 
  mutate(x = -x0, y = -y0, z = z0 * z_scale) %>% 
  select(x,y,z)

plot_ridge <- dem_subset %>%
  render_ridge(n_ridges = 300, n_threshold = 0, z_shift = 5, z_threshold = 7) 

ggsave(
  plot_ridge, file = "R/figures/gis/ridge_loudenvielle_00.png",
  dpi=300, width=format$a4[1], height = format$a4[2],
  scale=1, units="mm")

```

