# use open-data digital elevation model (ESRI ASCII Raster)

```{r setup}
library(tidyverse)
library(stars)

knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)

# functions
source("R/functions_geometry.r")
source("R/functions_gis.r")
```

## Outline

The main idea is to convert 3D point sets to 2D lines. Digital Elevation Model (DEM) represents elevation as function of latitude and longitude. These illustrations are made of multiple lines of elevation as a function of longitude, for discrete latitude values. Overplotting is avoided (hidden lines) by checking for a minimal distance between two lines. These three examples explore the use of :

* a coarse and easy source for elevation data ([elevatr](https://github.com/jhollist/elevatr) package), for a broad view (~ 450 km)
* a precise DEM at a the scale of a specific valley (Ossau, FR, ~ 20 km) using IGN 5m RGE ALTI DEM ([link](https://geoservices.ign.fr/documentation/diffusion/telechargement-donnees-libres.html#rge-alti-5-m)) 
* a focus on a narrow zone (Loudenvielle, FR, ~ 3 km), with a fine initial sampling resolution for the elevation (30 cm, as reported [here](https://www.geoportail.gouv.fr/carte?c=2.4206324117987448,48.81510668542717&z=10&l0=GEOGRAPHICALGRIDSYSTEMS.MAPS.SCAN-EXPRESS.STANDARD::GEOPORTAIL:OGC:WMTS(1)&l1=ELEVATIONGRIDCOVERAGE.HIGHRES.QUALITY::GEOPORTAIL:OGC:WMTS(1)&permalink=yes)). 


```{r plot_pyrenees, eval=FALSE}

# Large scale ridge plot with open-access data
# define location and bounding box
coord = c(lat=43.231196, lon = 0) # Tarbes
z_scale = 2 # ratio of elevation:cell unit

# convert DEM to a dataframe and subset based on bounding box
point <- data.frame(lon = coord["lon"], lat = coord["lat"]) %>% 
  st_as_sf(coords = c("lon", "lat")) %>% st_set_crs(4326) %>% st_transform(crs = 2154) 

box <- buffer_rectangle(
  point, size = 450E3, y_shift = - 60E3, x_shift = 30E3,
  orientation="h", ratio = 2)

# use AWS terrain tile (30m)
dem <- elevatr::get_elev_raster(point, z = 7) 

dem_subset <- dem %>% st_as_stars() %>% 
  st_crop(box) %>% as_tibble() %>%
  select(x,y, z = 3) %>% 
  mutate(x = x, y = y, z = z * z_scale)

data_ridge <- dem_subset %>%
  render_ridge(n_ridges = 200, z_shift = 1000, z_threshold = 100) %>% 
  filter_ridge(length_n = 5)

# plot
plot_ridge <- data_ridge %>%
  ggplot(aes(xn, zn, group = y_rank)) +
  geom_line(alpha = 0.6, size = 0.2) +
  coord_fixed() + theme_void()

ggsave(
  plot_ridge, file = "R/figures/gis/ridge_pyrenee_00.png",
  dpi=300, width=format$a3[2], height = format$a3[2]/ratio$uni ,
  scale=1, units="mm")

```


```{r plot_ossau, eval=FALSE}

# Valley scale, French DEM source (5m, 30m source)
# define location and bounding box
coord = c(lat=42.9136778, lon=-0.4015139) # Cezy

# use RGE ALTI 5m DEM
dem <- coord %>% read_dem(buffer = 12000, dep="64", source = "rge_alti")

# convert DEM to a dataframe and subset based on bounding box
point <- data.frame(lon = coord["lon"], lat = coord["lat"]) %>% 
  st_as_sf(coords = c("lon", "lat")) %>% st_set_crs(4326) %>% st_transform(crs = 2154) 

box <- buffer_rectangle(point, size = 22E3, y_shift = 1E3, orientation="v")

# rotate the DEM matrix to draw ridgelines from north to south
dem_subset <- dem %>% st_crop(box) %>% 
  as_tibble() %>% mutate(x = -x, y = -y)

data_ridge <- dem_subset %>%
  render_ridge(n_ridges = 300, n_drop = 10, z_shift = 10, z_threshold = 8) %>% 
  filter_ridge(length_n = 15)

# plot
plot_ridge <- data_ridge %>%
  ggplot(aes(xn, zn, group = y_rank)) +
  geom_line(alpha = 0.6, size = 0.2) +
  coord_fixed() + theme_void()

ggsave(
  plot_ridge, file = "R/figures/gis/ridge_ossau_01.png",
  dpi=300, width=format$a3[2], height=format$a3[2]/ratio$wide, scale=1, units="mm")

```


```{r plot_loudenvielle, eval=FALSE}

# Ridge line rendering with true HR DEM data (5m, 30cm source)

coord = c(lat=42.744617, lon = 0.410614) # Loudenvielle
z_scale = 1 # ratio of elevation:cell unit

# convert DEM to a dataframe and subset based on bounding box
point <- data.frame(lon = coord["lon"], lat = coord["lat"]) %>% 
  st_as_sf(coords = c("lon", "lat")) %>% st_set_crs(4326) %>% st_transform(crs = 2154) 

# RGE ALTI (5m)
dem <- coord %>% read_dem(buffer = 5000, dep="65", source = "rge_alti")

box <- buffer_rectangle(point, size = 3E3, orientation="v")

# rotate the DEM matrix to draw ridgelines from north to south
dem_subset <- dem %>% st_as_stars() %>% 
  st_crop(box) %>% as_tibble() %>%
  select(x0=x, y0=y, z0 = 3) %>% 
  mutate(x = -x0, y = -y0, z = z0 * z_scale) %>% 
  select(x,y,z)

data_ridge <- dem_subset %>%
  render_ridge(n_ridges = 300, z_shift = 5, z_threshold = 7) 

# plot
plot_ridge <- data_ridge %>%
  ggplot(aes(xn, zn, group = y_rank)) +
  geom_line(alpha = 0.6, size = 0.2) +
  coord_fixed() + theme_void()

ggsave(
  plot_ridge, file = "R/figures/gis/ridge_loudenvielle_00.png",
  dpi=300, width=format$a4[1], height = format$a4[2],
  scale=1, units="mm")

```


```{r plot_heric, eval=FALSE}

# Use full resolution DEM data (1 line / 5 m)

coord = c(lat=43.571297, lon = 2.970772) # Heric
z_scale = 1 # ratio of elevation:cell unit

# convert DEM to a dataframe and subset based on bounding box
point <- data.frame(lon = coord["lon"], lat = coord["lat"]) %>% 
  st_as_sf(coords = c("lon", "lat")) %>% st_set_crs(4326) %>% st_transform(crs = 2154) 

# RGE ALTI (5m)
dem <- coord %>% read_dem(buffer = 5000, dep="34", source = "rge_alti")

box <- buffer_rectangle(
  point, size = 4000, y_shift = 1000, x_shift = 0,
  orientation="v")

dem_subset <- dem %>% st_as_stars() %>% 
  st_crop(box) %>% as_tibble() %>%
  select(x0=x, y0=y, z0 = 3) %>% 
  mutate(x = x0, y = y0, z = z0 * z_scale) %>% 
  select(x,y,z)

data_ridge <- dem_subset %>%
  render_ridge(n_ridges = 0, n_lag = 500, z_shift = 0, z_threshold = 2) %>% 
  filter_ridge(length_n = 2, length_x = 2/100, dist_y = 0.90) 

# plot
plot_ridge <- data_ridge %>%
  ggplot(aes(xn, zn)) +
  geom_line(aes(group = y_rank), size = 0.2) +
  geom_line(
    data = data_ridge %>% group_by(xn) %>% slice_max(zn, n=1),
    size = 0.2) +
  lims(x=c(0, 2000), y=c(250,1100)) +
  coord_fixed() + theme_void()

file <- "ridge_heric_00"

ggsave(
  plot_ridge, file = glue::glue("R/figures/ode/{file}.png"),
  dpi=300, width=format$a4[2], height = format$a4[1],
  scale=1, units="mm")

ggsave(
  plot_ridge, file = glue::glue("R/figures/ode/{file}.svg"),
  width=format$a4[2], height = format$a4[1],
  scale=1, units="mm")

glue::glue("vpype read {file}.svg linesimplify linemerge --tolerance 0.5mm  linesort  write {file}_plot.svg")
```

```{r plot_arves, eval=FALSE}

# Valley scale, High mean elevation, French DEM source (5m, 1m source)

# define location and bounding box
coord = c(lat=45.174716, lon = 6.335077) # Aiguilles d'Arves 

z_scale = 1 # ratio of elevation:cell unit

# convert DEM to a dataframe and subset based on bounding box
point <- data.frame(lon = coord["lon"], lat = coord["lat"]) %>% 
  st_as_sf(coords = c("lon", "lat")) %>% st_set_crs(4326) %>% st_transform(crs = 2154) 

# RGE ALTI (5m)
dem <- coord %>% read_dem(buffer = 8000, dep="73", source = "rge_alti")

box <- buffer_rectangle(
  point, size = 8000, y_shift = -3500, x_shift = 0,
  orientation="v")

dem_subset <- dem %>% st_as_stars() %>% 
  st_crop(box) %>% as_tibble() %>%
  select(x0=x, y0=y, z0 = 3) %>% 
  mutate(x = - x0, y = - y0, z = z0 * z_scale) %>% 
  select(x,y,z)

data_ridge <- dem_subset %>%
  render_ridge(n_ridges = 800, n_drop = 1, n_lag = 500, z_shift = 0, z_threshold = 4) %>% 
  filter_ridge(length_n = 2) 

# plot
plot_ridge <- data_ridge %>%
  ggplot(aes(xn, zn)) +
  geom_line(aes(group = y_rank), size = 0.2) +
  geom_line(
    data = data_ridge %>% group_by(xn) %>% slice_max(zn, n=1),
    size = 0.2) +
  #lims(x=c(0, 6000), y=c(2500, 7000)) +
  coord_fixed() + theme_void()

file <- "ridge_arves_00"

ggsave(
  plot_ridge, file = glue::glue("R/figures/ode/{file}.png"),
  dpi=300, width=format$a3[2], height = format$a3[1],
  scale=1, units="mm")

ggsave(
  plot_ridge, file = glue::glue("R/figures/ode/{file}.svg"),
  width=format$a3[2], height = format$a3[1],
  scale=1, units="mm")

glue::glue("vpype read {file}.svg linesimplify linesort  write {file}_plot.svg")

```
