# Explore the use of sequences to generate plant-like visualizations

```{r setup}
library(tidyverse)
library(ggforce)
library(furrr)
library(cowplot)

plan(multisession)

knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)

# functions
source("R/paper_functions.r")
source("R/sequence_functions.r")

```

## Outline

A sequence of integers is generated from a set of rules (here, the Collatz [conjecture](https://en.wikipedia.org/wiki/Collatz_conjecture)). Then, a curve is defined by mapping sequence elements to segment length, with a fixed angle between segments. Multiple curves are computed to constitute a node, and multiple nodes to constitute a stem. The only random elements are initial value for the sequence, and the angle between curve segments. These two examples explore the use of :

* generating multiple curves and shifting vertically their starting point by a fixed amount (plot_node_dense)
* generating multiple node with attributes (number of leaves, scale, angle, position) as a function of a stem-like structure (plot_stem).

```{r plot_node_dense, eval=FALSE}
# plot a single node with dense polygons

p_seed = 5
p_node = 300
p_imax = 100
p_lmax = 3000
p_angle = 40
p_shift = 4
p_width = 10
p_shape = "spline"

# seeds 4, 8, 11
data_dense <- gen_node(
  n = p_node, imax = p_imax, lmax = p_lmax,
  amin = -p_angle, amax = p_angle, 
  shift = p_shift, width = p_width, seed = p_seed, shape=p_shape)

plot_dense <- data_dense %>%
  render_node(radius = 0, margin = 10,
    xlim = c(-100, 100), ylim = c(180,460))

ggsave(
  plot_dense, file = "R/figures/sequence/collatz_dense_05.png",
  width=format$a4[1], height = format$a4[2],
  dpi=300, scale=1, units="mm")

```



```{r plot_stem, eval=FALSE}

set.seed(1)

p_stem = 73 # starting point for the stem sequence
p_stem_angle = 10 # angle between successive stem segments
p_node = 35 # node number in the branch
p_node_angle = c(140, 10) # mean and sd of angle between nodes and stem
p_scale = 0.5 # stem scale

# set node parameters along the stem
data_topology <- tibble(
  node = seq_len(p_node),
  leaf = seq(15, 10, len = p_node),
  angle = seq(10, 5, len = p_node),
  lmax = seq(700, 400, len = p_node),
  scale = seq(1.2, 0.3, len = p_node)
  )

# generate nodes as a function of parameter list
data_node <- data_topology %>% 
  mutate(
    data = pmap(
      list(leaf, angle, lmax, scale),
      ~ gen_node(
        n = ..1, amin = -..2, amax = ..2, lmax = ..3, scale = ..4,
        width = 20))
    )
# generate stem geometry and node transformation parameters
data_stem <- gen_leaf(p_stem, a=p_stem_angle) %>% slice(1:p_node) %>% 
  mutate(across(x:yend, ~ . * p_scale)) %>% 
  mutate(
    n = rev(n),
    a = seq_alt(n=p_node, m=p_node_angle[1], sd=p_node_angle[2]),
    a = (a + angle) * pi/180) 

# merge stem and node data
data_plot <- data_node %>% 
  left_join(data_stem %>% select(node=n, x0=x, y0=y, a)) %>% 
  mutate(data_t = pmap(list(data, x0, y0, a), ~ r_t(..1, ..2, ..3, ..4)))

# plot
p <- ggplot() +
  geom_shape(
    data = data_stem %>% transform_path(width = 15),
    aes(x, y), alpha=0.2, size=1,
    radius = unit(2.5, 'pt'), expand = unit(2.5, 'pt'), 
    fill="white", color="darkgrey") +
  geom_path(data = data_stem, aes(x=x,y=y), color="darkgrey", size = 0.5) +
  geom_shape(
    data = data_plot %>% unnest(data_t),
    aes(x,y, group = interaction(node, id)),
    color="darkgrey", fill="white", size = 0.5) + 
  coord_fixed() + theme_void()


# export
file <- "collatz_fern_01"

ggsave(
  p, file = glue::glue("R/figures/sequence/{file}.png"),
  width=format$a4[2], height = format$a4[1],
  dpi=400, scale=1, units="mm")

# cmd <- glue::glue("vpype  read {file}_traced.svg  filter --min-length 1mm  linemerge --tolerance 0.5mm  linesort  write {file}_plot.svg ")

```




