# Generative text system based on splines curves

```{r setup}
library(tidyverse)
library(ggforce)
library(ggraph)
library(tidygraph)
library(cowplot)

knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)

# function
source("R/splines_functions.r")

# parameters

# define coordinates systems for plots
coord_square <- coord_fixed(
  ratio = 1, expand = TRUE,
  xlim = c(-1, 1), ylim = c(-1, 1))

coord_ellipse <- coord_fixed(
  ratio = 1, expand = TRUE,
  xlim = c(-0.7, 0.7), ylim = c(-0.9, 0.9))   
```

## Outline

Glyphs are generated from splines defined by few control point sampled from a 2D space (square, ellipse). Glyphs are then mapped to characters to generate paragraphs with sensible text-like aesthetics. Tested alterations from this base system :

* draw glyphs from simple layouts independently from each other to emulate a script-like writing (plot_script).
* concatenate few layout to generate individual glyphs (plot_script_merge).
* concatenate glyphs to form word to emulate a cursive-like writing (plot_cursive).
* increase the size and number of control points to generate more detailed and complex shapes (plot_cursive_complex)


```{r plot_glyphs, eval=FALSE}
# generate glyph type by using splines
# https://inconvergent.net/2017/spline-script/

set.seed(1)
n_points = 6

data_glyphs <- tibble(pattern=1:100) %>% 
  mutate(
    layout = map(pattern, ~ layout_ellipse(n = n_points)),
    plot = map(layout, render_spline, coord = coord_ellipse)
  )

plot_glyphs <- plot_grid(plotlist = data_glyphs$plot, ncol=10, scale=0.8) + 
  theme(plot.margin = unit(c(1,1,1,1), "cm")) 

ggsave(
  plot_glyphs,
  file="./R/figures/splines/set_glyphs.png",
  dpi=300, width=210, height=210, scale=1, units="mm")

```


```{r plot_script, eval=FALSE}
# map glyphs to letters and generate text as glyphs sequences
# alterate individual glyphs with random variations

set.seed(1)
n_points = 6    # number of control points per character
n_col = 60      # number of characters per line (default 70)

# define a character map 
data_map <- tibble(
    pattern=1:29,
    character = c(letters[1:26], ".", ",", "?")
)

# add glyph variability using random noise on layout coordinates (10 reps)
data_glyphs <- data_map %>% 
  mutate(
    layout = map(pattern, ~ layout_ellipse(n = n_points))
  ) %>%  
  crossing(variation=1:10) %>% 
  mutate(
    layout_rng = map(
      layout,
      ~ unnest(.x, cols = c(x,y)) %>%
        mutate(across(x:y, jitter, amount=1/5))),
    plot = map(layout_rng, render_spline, width=0.3, coord=coord_ellipse),
    plot_bold = map(layout_rng, render_spline, width=0.5, coord=coord_ellipse)
  )

# generate a nonsense text
set.seed(10)
seq_text <- stringi::stri_rand_lipsum(n_paragraphs = 2)
seq_prop <- str_length(seq_text)/sum(str_length(seq_text))

# render each paragraph using a random variation for each character
data_script <- tibble(text = seq_text) %>% 
  mutate(plot = map(text, ~ render_script(., data=data_glyphs, ncol=n_col, scale=1.8)))

# export text
plot_script <- plot_grid(
  plotlist = data_script$plot,
  ncol=1, rel_heights=c(seq_prop)) +
  theme(plot.margin = unit(c(2,2,2,2), "cm"))

ggsave(
  plot_script,
  file="./R/figures/splines/text_script_00.png",
  dpi=300, width=210, height=210, scale=1, units="mm") 

# export map
plot_glyphs <- plot_grid(
  plotlist = data_glyphs[1:100,]$plot_bold, ncol=10, scale=0.8) +
  theme(plot.margin = unit(c(1,1,1,1), "cm"))

ggsave(
  plot_glyphs,
  file="./R/figures/splines/set_glyphs_variability.png",
  dpi=300, width=210, height=210, scale=1, units="mm")

```

```{r plot_script_merge, eval=FALSE}
# generate more complex shapes by concatenating n individual layouts

sample_letters <- function(n = 3) {
  sample(letters, n, replace = TRUE) %>% paste0(collapse = "") 
}

set.seed(1)

n_letters = 26  # number of letters in alphabet
n_tall = 4      # number of tall letters
size_tall = 8   # size of tall letters (units)
n_points = 4    # number of control points per character
n_chr = 3       # number of elements to be concatenated as new glyphs
n_col = 40      # number of characters per line (default 70)
n_length = 1    # total length of text (number of paragraph)

# define a character map 
data_map <- gen_charmap(n_letters, n_points, n_tall, size_tall)

data_glyphs <- data_map %>% 
  mutate(
    variation = 1,
    word = rerun(29, sample_letters(n = n_chr)),
    layout_word = map(word, ~ layout_word(., data_map)),
    plot = map(layout_word, ~ render_spline(., n = 300, coord = coord_fixed()))
    )

# generate a nonsense text
set.seed(10)
seq_text <- stringi::stri_rand_lipsum(n_paragraphs = n_length)
seq_prop <- str_length(seq_text)/sum(str_length(seq_text))

# render each paragraph concatenated glyphs
data_script <- tibble(text = seq_text) %>% 
  mutate(plot = map(text, ~ render_script(., data=data_glyphs, ncol=n_col, scale=1)))

# export text
plot_script <- plot_grid(
  plotlist = data_script$plot,
  ncol=1, rel_heights=c(seq_prop)) +
  theme(plot.margin = unit(c(2,2,3,2), "cm"))


# plot_glyph <- plot_grid(plotlist = data_glyphs$plot, scale = 0.9)

ggsave(
  plot_script,
  file="./R/figures/splines/text_script_01.png",
  dpi=300, width=210, height=210, scale=1, units="mm") 

```


```{r plot_cursive, eval=FALSE}
# generate words by concatenating individual graphs layouts

# parameters
set.seed(1)

n_letters = 26  # number of letters in alphabet
n_tall = 5      # number of tall letters
size_tall = 5   # size of tall letters (units)
n_points = 4    # number of control points per character
n_col = 70      # number of characters per line (default 70)
n_length = 3    # total length of text (number of paragraph)

# define a character map and randomly modify few characters in y scaling
data_glyphs <- gen_charmap(n_letters, n_points, n_tall, size_tall)

# generate a nonsense text
set.seed(10)
seq_text <- stringi::stri_rand_lipsum(n_paragraphs = n_length)

# generate words from sequences of random length
data_cursive <- tibble(text = seq_text) %>% 
  mutate(line = round(str_replace_all(text, " ", "") %>% str_length() / n_col)) %>% 
  mutate(plot = map(
    text,
    ~ render_cursive(., data = data_glyphs, ncol = n_col, scale = 0.8)))

# export text
plot_cursive <- plot_grid(
  plotlist = data_cursive$plot,
  ncol=1, rel_heights = data_cursive$line) +
  theme(plot.margin = unit(c(2,1,3,0), "cm"))

ggsave(
  plot_cursive,
  file="./R/figures/splines/text_cursive_00.png",
  dpi=300, width=210, height=210, scale=1, units="mm")

```

```{r plot_cursive_large, eval=FALSE}
# generate longer and larger words with more detailed and complex shapes

# parameters
set.seed(1)

n_letters = 26  # number of letters in alphabet
n_tall = 5      # number of tall letters
size_tall = 5   # size of tall letters (units)
n_points = 4    # number of control points per character
n_col = 40      # number of characters per line (default 70)
n_length = 1    # total length of text (number of paragraph)

# define a character map and randomly modify few characters in y scaling
data_glyphs <- gen_charmap(n_letters, n_points, n_tall, size_tall)

# generate a nonsense text
set.seed(10)
seq_text <- stringi::stri_rand_lipsum(n_paragraphs = n_length)

# generate words from sequences of random length
data_cursive <- tibble(text = seq_text) %>% 
  mutate(line = round(str_replace_all(text, " ", "") %>% str_length() / n_col)) %>% 
  mutate(plot = map(
    text,
    ~ render_cursive(., data = data_glyphs, ncol = n_col, shift = 1, scale = 0.8)))

# export text
plot_cursive <- plot_grid(
  plotlist = data_cursive$plot,
  ncol=1, rel_heights = data_cursive$line) +
  theme(plot.margin = unit(c(2,1,4,1), "cm"))

ggsave(
  plot_cursive,
  file="./R/figures/splines/text_cursive_01.png",
  dpi=300, width=210, height=210, scale=1, units="mm")

```

```{r plot_cursive_long, eval=FALSE}
# generate longer and larger words with more detailed and complex shapes

# parameters
set.seed(8)

n_letters = 26   # number of letters in alphabet
n_tall = 5       # number of tall letters
size_tall = 20    # size of tall letters (units)
n_points = 5     # number of control points per character
size_seq = 10:20 # number of letters in generated words
n_words = 8     # number of generated words
n_col = 40       # number of characters per line

# define a character map and randomly modify few characters in y scaling
data_glyphs <- gen_charmap(n_letters, n_points, n_tall, size_tall)

# generate a fixed number of words with various length
set.seed(2)

seq_text <- stringi::stri_rand_strings(
  n = n_words, pattern = "[a-z]",
  length = sample(size_seq, size = n_words, replace = TRUE)) %>% 
  paste(., collapse = " ")

data_cursive <- tibble(text = seq_text) %>% 
  mutate(line = round(str_replace_all(text, " ", "") %>% str_length() / n_col)) %>% 
  mutate(plot = map(
    text,
    ~ render_cursive(., data = data_glyphs, ncol = n_col, n_points = 500, scale = 1)))

# plot word
plot_cursive <- plot_grid(
  plotlist = data_cursive$plot,
  ncol=1, rel_heights = data_cursive$line) +
  theme(plot.margin = unit(c(5,1,5,1), "cm"))

ggsave(
  plot_cursive,
  file="./R/figures/splines/text_cursive_02.png",
  dpi=300, width=210, height=210, scale=1, units="mm")

```

